---
title: "Prediction Assignment"
output: html_document
---

##Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

###loading libraries
```{r}
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
```

###setting the seed for reproduceability
```{r}
set.seed(1234)
```

###Loading the data set
```{r}
trainingset <- read.csv("C:/Users/abilkassymov/Desktop/pml-training.csv", na.strings=c("NA","#DIV/0!", ""))
testingset <- read.csv("C:/Users/abilkassymov/Desktop/pml-testing.csv", na.strings=c("NA","#DIV/0!", ""))
```

###Check dimensions
```{r}
dim(trainingset) 
dim(testingset)
```

###Delete missing values
```{r}
trainingset<-trainingset[,colSums(is.na(trainingset)) == 0]
testingset <-testingset[,colSums(is.na(testingset)) == 0]
trainingset   <-trainingset[,-c(1:7)]
testingset <-testingset[,-c(1:7)]
```

###Check dimensions
```{r}
dim(trainingset)
dim(testingset)
```
The training data set contains 53 variables and 19622 obs.
The testing data set contains 53 variables and 20 obs.

###Create data partition
```{r}
subsamples <- createDataPartition(y=trainingset$classe, p=0.75, list=FALSE)
subTraining <- trainingset[subsamples, ] 
subTesting <- trainingset[-subsamples, ]
dim(subTraining)
dim(subTesting)
```
the training data set is partionned into 2 sets: subTraining (75%) and subTest (25%)

###look at data
A plot of the outcome variable will allow us to see the frequency of each levels in the subTraining data set and compare one another.
```{r}
plot(subTraining$classe, col="blue", main="Bar Plot of levels of the variable classe within the subTraining data set", xlab="classe levels", ylab="Frequency")
```

###create first model
```{r}
model1 <- rpart(classe ~ ., data=subTraining, method="class")
```

###Predicting
```{r}
prediction1 <- predict(model1, subTesting, type = "class")
```

###Plot of the Decision Tree
```{r}
rpart.plot(model1, main="Classification Tree", extra=102, under=TRUE, faclen=0)
```

###Test results
```{r}
confusionMatrix(prediction1, subTesting$classe)
```
Random Forest algorithm performed better than Decision Trees

Random Forest model was 0.995 (95% CI: (0.993, 0.997))
Decision Tree model was 0.739 (95% CI: (0.727, 0.752))

Random Forest is better and the expected out-of-sample error is estimated at 0.5%

###create second model
```{r}
model2 <- randomForest(classe ~. , data=subTraining, method="class")
```

###Predicting
```{r}
prediction2 <- predict(model2, subTesting, type = "class")
```

###Test results on subTesting data set:
```{r}
confusionMatrix(prediction2, subTesting$classe)
```

###submission
```{r}
predictfinal <- predict(model2, testingset, type="class")
predictfinal

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(predictfinal)
```